---
title: "Topic Modeling"
date: "2024-03-26"
categories: [NLP]
editor_options: 
  chunk_output_type: inline
footnotes-hover: true
execute: 
  warning: false
---

```{r}
#| echo: false
library(DiagrammeR)
library(tidytext)
library(tidyverse)
grViz("
  digraph {
  layout = dot
  node [shape = rectangle,style=filled,fixedsize=False]
  edge[color=grey,arrowhead=vee]
  A0[label = 'Corpus']
  A[label = 'Pre-Processed Text']
  B[label = 'Text Vectorisation']
  B1[label = 'Dimensionality Reduction']
  C[label = 'Topic Modeling', fontname = 'times-bold']
  C1[label = 'LDA']
  C2[label = 'NMF']


  A0->A
  A->B
  rank=same{B->B1} 
  B->C
  C->{C1 C2}

  


  }
  "
  )
```

## The Process So Far:

```{r}
#| label: pre-process-data
library(tidyverse)
library(tidytext)
library(textstem)

# Input Dataset
newtest_df <- scriptuRs::new_testament %>% tibble()
matthew_df <- newtest_df %>% 
  filter(book_title == "Matthew")

# Identify Stop Words
more_stop_words <- tibble(word = c("thou", "thee", "thy", "hath", "shalt", "woe", "ye"), lexicon = "Az")

stop_words <- bind_rows(stop_words, more_stop_words)

# Tokenize Data and remove stop words
matthew_token_df <- matthew_df %>% 
  unnest_tokens(
    input = text,
    output = "word"
  ) %>% 
  anti_join(stop_words)

# Lemmatize words
matthew_lemma_df <- matthew_token_df %>% 
  mutate(lemma = lemmatize_words(word))

# TDM Matrix (with TF-IDF weights)
matthew_dtm <- matthew_lemma_df %>% 
  count(chapter_number, lemma) %>% 
  cast_dtm(document = chapter_number,
           term = lemma,
           value = n,
           weighting = tm::weightTfIdf)
```

## Topic Modeling

Topic modeling is a process that aims to discover the underlying themes (or **topics**) in a corpus. The goal is to use the word distribution within documents to identify latent topics without pre-defined labels or categories.

In our example, we are using topic modeling as we are interested in identifying the themes that are present in the book of Matthew.

Topic modeling uses an algorithm to extract the topics based from the text vector that we previously created. The most common algorithm (and the one we will focus on today) is **Latent Dirichlet Allocation (LDA)**. However, another common algorithm is **Non-negative Matrix Factorization (NMF)**, which is used mostly when there is not enough computing power to perform LDA.

### Latent Dirichlet Allocation

A useful way to understand LDA is to breakdown its name: **Latent** refers to the fact that the topics in the text are not directly observable. For anyone familiar with other "Latent" models, the concept is the same. By using connections between elements in the observable data, we can identify the latent topics that these connections refer to. For example, if a document frequently mentions "dogs" and "cats", the latent variable might be "pets".

**Dirichlet** refers to the specific probability distribution (a dirichlet distribution) that represents the document-topic and topic-word distributions. To understand dirichlet allocation requires a solid understanding of probability distributions and is out of the scope of this series. The **Allocation** part of LDA refers to the process of allocating words to topics and topics to documents based on the distributions from the model.

The <code>LDA()</code> function from the **{topicmodels}** package will do the heavy-lifting for us in performing our LDA.

-   we need to convert our TF-IDF back to just term frequencies

-   We do TF-IDF to help us get familiar with the data and to inform us of any dimensionality reduction processes that need to take place.

```{r}
library(topicmodels)

# convert back to term frequency

matthew_tf <- matthew_lemma_df %>% 
  count(chapter_number, lemma) %>% 
  cast_dtm(document = chapter_number,
           term = lemma,
           value = n,
           weighting = tm::weightTf)



lda_model <- LDA(matthew_tf, k = 5, control = list(seed = 1234))

terms(lda_model)
```
